---
title: "Viz and Eda1"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


Let's make a plot!

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```


# Pipes and stuff
useful to set up as pipe so you can do other functions along with it

```{r}
ggp_nyc_weather = 
  weather_df |> 
  filter(name =="CentralPark_NY") |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_nyc_weather
```

# Fancy plot
color applie to points and smooth curve
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth()
```

color applies just to points and not to smooth
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) +
  geom_smooth(se = FALSE)
```
se false to get rid of error bars

use alpha shading to make points more transparent; to better see all data points

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name, alpha = 0.3)) +
  geom_smooth(se = FALSE)
```

Plot with facets
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

automatically places graphs in alaphabetical order; can alter this

diff fancy plot
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp, alpha = 0.3)) +
  geom_smooth() +
  facet_grid(. ~ name)
```

assignment of specific colors; add to component it applies to; don't put on ggplot line bc it thinks there's a variable called blue
do this
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(color = "blue")

```

not this
```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x = date, y = tmax, color = "blue")) +
  geom_point()
```

Jeff rarely assigns specific colors
adjusts transparency and size manually 

can use geom_hex if too many points are overlaid
kind of like a density plot

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

line plot
can be helpful when you have repeated measures
```{r}
weather_df |> 
  filter(name == "Molokai_HI") |> 
  ggplot(aes(x = date, y = tmax)) +
  geom_line() +
  geom_point()
```


# Univariate plotting
Histogram
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge")
```

to compare across groups, Jeff recommends a density plot (a smoothed over histogram)
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) +
  geom_density(alpha = 0.3, adjust = 2)
```

defaults are usually pretty good
adjust smooths over
double checking can be helpful

boxplots

```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_boxplot()
```

violin plots
(essentially make density plot vertical then you mirror it)
```{r}
ggplot(weather_df, aes(y = tmax, x = name)) +
  geom_violin()
```

ridge plot
creates 3 separate densites stacked onto each other vertically; helpful when you many groups to compare 
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) +
  geom_density_ridges()
```







