---
title: "Viz and Eda2"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

This results in a dataframe with `r nrow(weather_df)` observations on six variables.


## Plot 1. Labeling! 

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp (Degrees C)",
    color = "Location",
    caption = "Max vs mnin daily temp in three locations; data from rnoaa"
  )

```

# Scales! Helps to control tick marks

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp (Degrees C)",
    color = "Location",
    caption = "Max vs mnin daily temp in three locations; data from rnoaa"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15") +
  scale_y_continuous(
    position = "right",
    trans = "sqrt")
  )
```

You don't have to do the transformation as part of the plotting, You can do it as a step in cleaning your data.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp (Degrees C)",
    color = "Location",
    caption = "Max vs mnin daily temp in three locations; data from rnoaa"
  ) +
  scale_x_continuous(
    breaks = c(-15, 0, 15),
    labels = c("-15 C", "0", "15")) +
  scale_y_continuous(
    position = "right",
    limits = c(20,30))
```

Better to set limits when clenaing data instead of adding it here.

# Setting color

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp (Degrees C)",
    color = "Location",
    caption = "Max vs mnin daily temp in three locations; data from rnoaa"
  ) +
  scale_color_hue(h = c(100, 300))
 
```


```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp (Degrees C)",
    color = "Location",
    caption = "Max vs mnin daily temp in three locations; data from rnoaa"
  ) +
  viridis::scale_color_viridis(discrete = TRUE)
 
```

viridis works best for color_blindness and printing in gray scale. Best to use viridis color palette.

# Themes (just controlling thematic stuff; mnoving things around)
```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  labs(
    title = "Temperature plot",
    x = "Min daily temp (Degrees C)",
    y = "Max daily temp (Degrees C)",
    color = "Location",
    caption = "Max vs mnin daily temp in three locations; data from rnoaa"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_bw() +
  theme(legend.position = "bottom")
  
 
```

Theme bw resets how the plot looks so you want to do more thematic adjustment after theme_bw()

theme_classic() - no gridlines; you should have gridlines
theme_minimal()

# Set preferences at the beginning for theme
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d


# Data argument
only on smooth line

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(aes(color=name)) +
  geom_smooth()
```

smooth line for each subset color name

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth()
```

can put a diff data frame on top of your graph. useful if you have to do a more complex summary

```{r}
nyc_weather_df <-
  weather_df |> 
  filter(name == "CentralPark_NY")

hawaii_weather_df <-
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(nyc_weather_df, aes(x = date, y = tmax)) +
  geom_point() +
  geom_line(data = hawaii_weather_df)
```


# facet
facet_grid(row ~ columns)

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  facet_grid(. ~ name)
```

# patchwork; package puts ggplot graphs next to each other
```{r}
ggp_temp_scatter <-
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")


ggp_prcp_density =
  weather_df |> 
  filter(prcp > 25) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5) +
  theme(legend.position = "none")

  
ggp_temp_scatter + ggp_prcp_density

ggp_tmax_date =
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom")

ggp_temp_scatter + ggp_prcp_density / ggp_tmax_date


```


## data manipulation
name variable is a character they are strings
factor variables are categorical variabes that have an order
this is important bc ggplot has to figure out on x axis and order of colors; every time you need to plot characters ggplot makes it factor but orders according to alphabetical order; in order to get a diff order have to do the following steps
```{r}
weather_df |> 
  mutate(
    name = fct_relevel(name, c("Molokai_HI"), "CentralPark_NY", "Waterhole_WA")
  ) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()

weather_df |> 
  mutate(
    name = fct_reorder(name, tmax)
  ) |> 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

# complicated FAS plot
sketch out plot you want to make first on paper. figure out how to tidy appropriately 

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

pups_df = 
  read_csv("data/FAS_pups.csv") |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"))

fas_df = 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    values_to = "pn_day"
  ) |> 
  drop_na() |> 
  mutate(outcome = fct_reorder(outcome, pn_day)) |> 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_violin() +
  facet_grid(day_of_tx ~ outcome)
  
```











